---
title: "Tennis Player Survival Analysis with Shiny"
runtime: shiny
output: html_document
---
```{r , echo = FALSE, warning = FALSE, message = FALSE, results='hide'}
library(data.table)
library(purrr)
library(ggplot2)
library(caret)
library(dplyr)
full_data <- read.csv("ATP.csv", stringsAsFactors = FALSE)
#Convertion of the variables is copied and pasted by Sijo VM, thanks!
number_columns <- c("winner_ht", "winner_age","loser_ht","loser_age" )
factor_columns <- c("tourney_id","surface", "draw_size", "tourney_level", "winner_id", "winner_entry",
                    "winner_hand", "winner_ioc", "loser_id", "loser_entry","loser_hand",
                    "loser_ioc","best_of","round") 

integer_columns <- c("match_num" ,"winner_seed","winner_rank","winner_rank_points","loser_seed",
                     "loser_rank", "loser_rank_points","minutes", "w_ace","w_df", "w_svpt","w_1stIn",
                     "w_1stWon","w_2ndWon","w_SvGms", "w_bpSaved", "w_bpFaced","l_ace","l_df","l_svpt",
                     "l_1stIn","l_1stWon","l_2ndWon","l_SvGms", "l_bpSaved","l_bpFaced")
character_columns <- c("tourney_name", "tourney_date","winner_name", "loser_name", "score")

for(i in number_columns){
  full_data[,i] <- as.numeric(full_data[,i])
} 
for(i in factor_columns){
  full_data[,i] <- as.factor(full_data[,i])
}  
for(i in integer_columns){
  full_data[,i] <- as.integer(full_data[,i])
}  
for(i in character_columns){
  full_data[,i] <- as.character(full_data[,i])
}


full <- as.data.table(full_data)
When_did_Roger_win <- function(Name){
#When I like to improve the function, I uncomment the player with the most wins:
#Name = "Roger Federer" 

Player <- full[full$winner_name == Name| full$loser_name == Name]
setkey(Player, winner_name)
as.character(Player$winner_name)
#If the chosen player is winner_name, he gets a 1, if not, he gets a 0.
Player[Name, survival:= 1]
Player[!Name, survival:= 0]

#If either the winner or the loser, or both are not seeded, a 0 is inserted.
Player$winner_seed[is.na(Player$winner_seed)]<-0
Player$loser_seed[is.na(Player$loser_seed)]<-0

#If the winner has a better seed then the loser, a 1 is inserted
Player[winner_seed < loser_seed & loser_seed != 0 , better_seed:= 1]
Player[winner_seed >= loser_seed, better_seed:= 0]

Player[as.character(winner_entry) == as.character(loser_entry), same_entry:= 1]
Player[as.character(winner_entry) != as.character(loser_entry), same_entry:= 0]

#Does size matter? If the winner is taller than the loser, we insert a 1.
Player$winner_ht[is.na(Player$winner_ht)]<-0
Player$loser_ht[is.na(Player$loser_ht)]<-0
Player[winner_ht > loser_ht , taller:= 1]
Player[winner_ht <= loser_ht , taller:= 0]

#Does the playing hand matter for the survival analysis?
Player[as.character(winner_hand) == as.character(loser_hand), same_hand := 0]
Player[as.character(winner_hand) != as.character(loser_hand), same_hand := 1]
#Player[,table(is.na(same_hand))]
#table(is.na(Player))

#Do older players play better? 
Player$winner_age[is.na(Player$winner_age)]<-0
Player$loser_age[is.na(Player$loser_age)]<-0

Player[winner_age > loser_age & loser_age != 0, older:= 1]
Player[winner_age <= loser_age |loser_age == 0 | winner_age == 0, older:= 0]
#Player[,table(older)]
#table(is.na(Player))

#If the winner ranks in the overall ATP ranking before the loser, a 1 is inserted.

Player$winner_rank[is.na(Player$winner_rank)]<-0
Player$loser_rank[is.na(Player$loser_rank)]<-0
Player[winner_rank < loser_rank, better_ranked:= 1]
Player[winner_rank >= loser_rank | loser_rank ==0, better_ranked:= 0]


#This variable insert a 1 if the games take longer than the median,the minutes variable will be included as well.
Player$minutes[is.na(Player$minutes)]<-0.1
median_played <- median(Player$minutes, na.rm = TRUE)
Player[minutes > median_played , overtime:= 1]
Player[minutes <= median_played & minutes >2, overtime:= 0]
Player[minutes < 3, overtime:= 2]

#This is no modification to the best_of variable, but improves
best_of_three <- 3
best_of_five <- 5
Player[best_of == best_of_five, distance:= 1]
Player[best_of == best_of_three, distance:= 0]
Player[best_of == 0, distance:= 2]

variables<-Player[,.(survival, same_entry, taller, same_hand, older,
          better_ranked, surface, better_seed, distance)]

set.seed(1)
rows <- sample(nrow(variables))
variables<-variables[rows,]
split<-round(nrow(variables)* 0.6)
train <- variables[1:split,]
test <- variables[(split+1):nrow(variables)] 
nrow(train) / nrow(variables)

model <- train(factor(survival)~. ,
               tuneLength = 3, #the more the costlier
               data = train, method = "ranger",
               importance = 'impurity', #Has to be included
               trControl = trainControl(method = "cv", number = 5, verboseIter = TRUE),
               classification = TRUE)
#plot(model)

result_test = predict(model, test, "raw")
#result_test
confu <- confusionMatrix(result_test, test$survival)
varimps <- varImp(model)
print(confu)

plot(varimps, main = Name)

}

hall_of_fame <- sort(table(full$winner_name),decreasing=TRUE)[1:50]
as.vector(hall_of_fame)
top_player <- names(hall_of_fame)
top_player
```
#Variable Importance Measure
The Variable Importance Measure shows which variables of opponent, surface and game duration matter in order to termine a victory. For the random forest approach only out-of-the-game variables are inclued, since they are known before the game. All variables are processed to `0` or `1` variables, except the `surface` variable.
The variables include:

* **better_ranked** Is the winner better ranked than the loser? 

* **older** Is the winner `older` than the loser?

* **surface** What surface has the match been played on: `Carpet`, `Hard`, `Grass`or `Clay`?

* **better_seed** Is the winner better ranked than the loser?

* **taller** Is the winner taller than the loser?

* **same_entry** Did any of the player entry via wildcard or qualification?

* **same_hand** Do the player's hand matter?

* **distance** Did the game take longer than the median?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r shiny, echo=FALSE}
inputPanel(
  selectInput("Players", label = "Number of bins:",
              choices = top_player, selected = "Roger Federer")
)

renderPlot({
  When_did_Roger_win(input$Players)
})
```
